hormones

### 2019-12-02 analysis/07_hormones.Rmd
new hormone analysis

```{r wrangle-prolactin}
prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1)

# keep only samples realated to parental care
prolactin <- prolactin %>% filter(Study %in% c("Baseline", "ParentalCare"))

# create lists of factos for characterization and manipluation studies

# rename some of the treatments levels, store in new columns
prolactin <- prolactin %>%
    mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           treatment = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc.d17" = "Inc_d17",
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "Inc_d3",
                             "inc.d3" = "inc_d3",
                            "inc.d9" = "Inc_d9",
                             "inc.d9" = "inc_d9",
                            "m.inc.d9" = "M_Inc9",
                            "m.inc.d9" = "M_inc9",
                            "m.inc.d3" = "M_Inc3",
                            "m.inc.d8" = "M_Inc8",
                            "m.inc.d8" = "M_inc8",
                            "m.inc.d17" = "M_Inc17",
                            "m.n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels1))

colnames(prolactin)[colnames(prolactin)=="Prolactin ng/mL"] <- "plasma_conc"
prolactin$hormone <- "prolactin"
prolactin <- prolactin %>% 
              select(study, treatment, sex, hormone, plasma_conc)  %>% 
              drop_na()

summary(prolactin)
```

```{r wrangle-steroids}
PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)

PETC <- PETC %>% select(stage, sex, hormone, plasma_conc)  %>%
                mutate(treatment = fct_recode(stage,
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "inc_d3",
                            "inc.d9" = "inc_d9",
                            "m.inc.d9" = "m_incd9",
                            "m.inc.d3" = "m_incd3",
                            "m.inc.d8" = "m_incd8",
                            "m.inc.d17" = "m_incd17",
                            "m.n2" = "m_hatch",
                            "control" = "stress_hpg")) %>%
                filter(treatment %in% alllevels2) %>%   
                mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
                       study = fct_collapse(treatment,
                                characterization = charlevels,
                                manipulation = maniplevels1)) %>% 
              drop_na() %>%  droplevels()  %>% 
              select(study, treatment, sex, hormone, plasma_conc)
summary(PETC)
```

@@

```{r sexsteroids}
hormonecharplot("cort", "CORT (ng/mL)")
hormonecharplot("progesterone", "PROG (ng/mL)") 
d1 <- hormonecharplot("estradiol", "E (ng/mL)")
d2 <- hormonecharplot("testosterone", "T (ng/mL)")
plot_grid(d1,d2, nrow = 1)

>>>>

```{r wrangle-prolactin}
prolactin <- read_excel("../results/Pigeon prolactin concentrations juil 2018.xlsx", sheet = 1) %>% 
  filter(Study %in% c("Baseline", "ParentalCare")) %>%
    dplyr::mutate(sex = fct_recode(Sex,
                            "female" = "f",
                            "male" = "m"),
           treatment = fct_recode(Treatment,
                            "hatch" = "Hatch",
                            "inc.d17" = "Inc_d17",
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "Inc_d3",
                             "inc.d3" = "inc_d3",
                            "inc.d9" = "Inc_d9",
                             "inc.d9" = "inc_d9",
                            "m.inc.d9" = "M_Inc9",
                            "m.inc.d9" = "M_inc9",
                            "m.inc.d3" = "M_Inc3",
                            "m.inc.d8" = "M_Inc8",
                            "m.inc.d8" = "M_inc8",
                            "m.inc.d17" = "M_Inc17",
                            "m.n2" = "M_hatch",
                            "control" = "baseline",
                            "n5" = "N5", 
                            "n9" = "N9"),
           study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels1)) %>%
          dplyr::rename("plasma_conc" = "Prolactin ng/mL") %>%
          mutate(bird_id = gsub("[[:punct:]]", "." , ColorBands)) %>% 
          dplyr::mutate(hormone = "prolactin") %>% 
          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc)  %>% 
          drop_na()
head(prolactin)

prolactin2 <- read_csv("../results/prolactin.csv") %>% 
          dplyr::rename("plasma_conc" = "prolactin_ng_mL",
                        "treatment" = "stage") %>% 
           dplyr::mutate(hormone = "prolactin") %>% 
           dplyr::mutate(treatment = fct_recode(treatment,
                            "m.n2" = "m.hatch",
                            "prolong" = "inc.prolong",
                            "extend" = "extend.hatch"),
                          study = fct_collapse(treatment,
                                 characterization = charlevels,
                                 manipulation = maniplevels1)) %>% 
          dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) 
head(prolactin2)

same <- inner_join(prolactin2, prolactin, by = "bird_id")

ggplot(same, aes(x = plasma_conc.x, y = plasma_conc.y, label = bird_id)) + geom_point() + geom_text()
```

```{r wrangle-steroids}
PETC <- read_excel("../results/parental_care_hormone_RIA_data_master.xlsx", sheet = 2)  %>% 
                dplyr::mutate(treatment = fct_recode(stage,
                            "inc.d17" = "inc_d17",
                            "inc.d3" = "inc_d3",
                            "inc.d9" = "inc_d9",
                            "m.inc.d9" = "m_incd9",
                            "m.inc.d3" = "m_incd3",
                            "m.inc.d8" = "m_incd8",
                            "m.inc.d17" = "m_incd17",
                            "m.n2" = "m_hatch",
                            "control" = "stress_hpg")) %>%
                dplyr::filter(treatment %in% alllevels2) %>%   
                dplyr::mutate(sex = fct_recode(sex,
                            "female" = "f",
                            "male" = "m"),
                       study = fct_collapse(treatment,
                                characterization = charlevels,
                                manipulation = maniplevels1)) %>% 
              dplyr::mutate(bird_id = gsub("[[:punct:]]", "." , band_combo)) %>% 
              dplyr::select(study, treatment, sex, bird_id, hormone, plasma_conc) %>% 
             drop_na() %>%  droplevels()  
head(PETC)

```{r correlations}
PETCP <- left_join(prolactin2, PETC, by = "bird_id") %>% drop_na()
PETCP$treatment.x <- factor(PETCP$treatment.x, levels = alllevels2)
PETCP$hormone.y

plothormonecorrelations <- function(myhormone, myylab){
  PETCP %>% filter(hormone.y == myhormone) %>%
  ggplot(aes(x = plasma_conc.x, y = plasma_conc.y, color = treatment.x )) +
  geom_point() + 
  geom_smooth(method = "lm", se = F) +
  labs(x = "Prolactin (ng/mL)", y = myylab) +
  scale_color_manual(values = colorscharmaip) +
  facet_wrap(~treatment.x, nrow = 3) +
  theme(legend.position = "none")
}

plothormonecorrelations("cort", "Corticosterone (ng/mL")
plothormonecorrelations("estradiol", "Estradiol (ng/mL")
plothormonecorrelations("progesterone", "Progesterone (ng/mL")
plothormonecorrelations("testosterone", "Testosterone (ng/mL")
```

@@

```{r sexsteroids}
hormonecharplot("cort", "CORT (ng/mL)")
hormonecharplot("progesterone", "PROG (ng/mL)") 
d1 <- hormonecharplot("estradiol", "E (ng/mL)")
d2 <- hormonecharplot("testosterone", "T (ng/mL)")
plot_grid(d1,d2, nrow = 1)
```
